---
title: "P8105 Homwork 2 Problem 2"
author: "Zanis Fang"
date: "9/26/2018"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**Read and clean the Mr. Trash Wheel sheet**

```{r load}
#load data
trash_wheel <- readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx",
																 sheet = 1,
																 range = readxl::cell_cols("A:N"))

trash_wheel <- trash_wheel %>% 
	             janitor::clean_names() %>% 
	             # omit rows counting totals, note only dumpsters have date information
	             filter(!is.na(date)) %>% 
	             # convert sports balls to integer variable
	             mutate(sports_balls = sports_balls %>% round(0) %>% as.integer()) %>% 
	             # note some NAs in "year" variable, simply get year from "date" variable
	             mutate(year = lubridate::year(lubridate::ymd(date)))

```
</br>
</br>

**Read and clean precipitation data from 2016 and 2017**

```{r precipitation}

# define a function to load data and clean data
prcp_load_clean <- function(wh_year) {
	                 
	                 # load data
	                 readxl::read_xlsx("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx",
		                      					sheet = paste(wh_year, "Precipitation"),
												      			range = "A2:B14"
														      	) %>%
		               
		               janitor::clean_names() %>%
		               # remove rows without data
	          	     filter(!is.na(total)) %>% 
		               # add year variable
		               mutate(year = wh_year)
	       
}

# combine two tables
prcp_data <- rbind(prcp_load_clean(2016), prcp_load_clean(2017))

# convert month to a character variable
prcp_data$month <- month.name[prcp_data$month]

```

In Mr. Trash Wheel data, it describes the amount of trash each dumpster colloects. There are `r nrow(trash_wheel)` individual dumpsters. In each dumpster, it categorizes the kind of trash, like "plastic bags", "cigarette butts", etc. For precipitation data in 2017, there are in total `r sum(subset(prcp_data, year == 2017, select = total))` inch precipitation. The median number of sports balls in a dumpster in 2016 is `r median(subset(trash_wheel, year == 2016, select = sports_balls)[[1]])`.